---
title: "climwin"
author: "Liam D. Bailey and Martijn van de Pol"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{climwin}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

`climwin` is a newly released package in R to help users choose a time period or 'window' over which to conduct ecological research. Although the general focus of the package will concern the impacts of climate, other novel applications can also be considered. This vignette will give an initial introduction to the package and its basic features. Future vignettes will provide more detail on the 'advanced' features present within the package.

************************************

## Why do we need climwin?

The characteristics of an organism are likely to change over time (e.g. body condition, behaviour), and these will influence the way in which an organism responds to its environment. As such, the time period or 'window' over which we choose to conduct ecological study can strongly influence the outcome and interpretation of our results. Yet there has been a tendency in ecological research to focus on a limited number of windows, often chosen arbitrarily. For example, there has been a strong focus on the biological impacts of average spring conditions. 

Without a critical comparison of different possible windows, we limit our ability to make meaningful conclusions from our research. If a biological trait displays no response to climate, it is difficult to determine if this is evidence of insensitivity to climate or whether the choice of time period is flawed. Even when we find a relationship between climate and the biological response, we cannot be sure that we have selected the period where the trait is most sensitive. Therefore, there is a need for a more effective method through which we can select our study period.

Realistically, manually testing and comparing all possible windows can be difficult and time consuming. With `climwin` we hope to overcome this problem by employing an exploratory approach to test and compare the effects of all possible windows. This will remove the need to arbitrarily select our study window, and will consequently improve our research outcomes and conclusions. Below we will outline the basic features of the climwin package, including how to carry out and interpret a climate window analysis.

****************************

## climatewin: The base function

`climatewin` is the main function of the climwin package. It uses an exploratory approach to investigate all possible windows and compares them using values of AICc.

### Under the hood

Let's imagine we have a dataset containing two variables: chick mass at hatching and date of measurement. Chick mass is likely to be measured on different dates due to biological differences (e.g. hatching date), therefore our dataset will look something like this:

| Date        | Mass (g) | 
|:------------|:---------|
| 04/06/2015  |    120   |
| 05/06/2015  |    123   |
| 07/06/2015  |    110   |
| 07/06/2014  |    140   |

We want to understand the relationship between temperature and chick mass, so we also have a second dataset containing daily temperature data.

| Date        | Temperature | 
|:------------|:------------|
| 01/06/2015  |    15       |
| 02/06/2015  |    16       |
| 03/06/2015  |    12       |
| 04/06/2014  |    18       |
| 05/06/2014  |    20       |
| 06/06/2014  |    23       |
| 07/06/2014  |    21       |

As is often the case, we have no prior knowledge to help us select the best climate window. To overcome this issue, `climatewin` will test all possible windows for us. In this case, we may start by testing the effect of average temperature 1-2 days before hatching. `climatewin` will calculate the corresponding average temperature for each biological record and create a new combined dataset.

| Date        | Mass (g) | Temperature [1 - 2 days] | 
|:------------|:---------|:-------------------------|
| 04/06/2015  |    120   |14                        |
| 05/06/2015  |    123   |15                        |
| 07/06/2015  |    110   |21.5                      |
| 07/06/2014  |    140   |21.5                      | 

Now that we have this new dataset, we can test the relationship between mass and temperature and determine an AICc value (Burnham, Anderson and Hayvaert 2011 [source](http://link.springer.com/article/10.1007/s00265-010-1029-6)). Let's use a basic linear model:

`lm(Mass ~ Temperature[1 - 2 days])`

We've tested the relationship between temperature and chick mass in one climate window, but we don't yet know how this window compares to others. `climatewin` will go back and carry out the same process on all other windows and compare the AICc value of each of these windows to that of a null model containing no climate. Our final outcome can be seen below:

|Window      |Model AICc |Null model AICc|Difference |
|:-----------|:----------|:--------------|:----------|
| 2 - 4 days | 1006      | 1026          | 20        |
| 1 - 3 days | 1015      | 1026          | 11        |
| 1 - 2 days | 1019      | 1026          | 7         |
| 4 - 5 days | 1020      | 1026          | 6         |
|...         |...        |...            |...        |

With this simple comparison, we can see that the strength of the window 2 - 4 days before hatching is not only better than a model with no climate (i.e. the value of AICc is smaller), but also better than other tested climate windows. In this case, we could conclude that chick mass was most strongly influenced by average temperature 2 - 4 days before hatching.

While this example is simplistic, it provides insight into the methodology behind `climatewin`. In practice, `climatewin` can be applied to large datasets testing thousands of possible windows, streamlining what would otherwise be a difficult and time consuming process.

### Getting started

Now that we understand how `climatewin` works, how do we go about using it? As with our simple example above, it is first necessary to create two separate datasets for the climate and biological data. These two datasets will be the basis for the climate window analysis.

To start using `climatewin` it's important to first understand the basic parameters. Below, we will discuss the parameters required to run a basic climate window analysis. More 'advanced' parameters will be discussed in additional vignettes.

For this example, we will use the `Mass` and `MassClimate` datasets from the `climwin` package.


**xvar**

To begin, we need to determine the predictor variable which we want to study. The parameter xvar is a list object containing our variable of interest (e.g. Temperature, Rainfall). While we will focus here on climate, it is possible to apply our `climatewin` methodology with non-climatic predictors.

`xvar = list(Temp = MassClimate$Temperature)`


**cdate/bdate**

Once we have established our predictor variable, we next need to tell `climatewin` the location of our date data. These two parameters contain the date variable for both the climate dataset (cdate) and biological dataset (bdate).

*NOTE: These date variables should be stored in a dd/mm/yyyy format*

`cdate = MassClimate$Date`

`bdate = Mass$Date`


**baseline**

The parameter baseline determines the model structure for the null model. The structure of the baseline model is highly versatile, with the potential to use multiple model types (lm, lmer, glm, glmer), include covariates and weights.

`baseline = lm(Mass ~ 1, data = Mass)`

Climate data from each tested climate window will be added as a predictor to this baseline model, to allow for a direct comparison between the null model and each climate window.


**cinterval**

Ideally, our predictive data will be available at a resolution of days. However, in some circumstances this may not be possible, especially when considering non-climatic data. With this in mind, cinterval allows users to adjust the resolution of analysis to days, weeks or months. Note that the choice of cinterval will impact our interpretation of parameters furthest and closest (see below). In our current example, daily data is available and so a resolution of days will be used.

`cinterval = "day"`



**range**

While our above example used a small dataset and tested a limited number of climate windows, the number of tested climate windows can be much larger, set by the parameter range. This parameter contains two value which set the upper and lower range for tested climate windows respectively. The values of range will correspond to the resolution chosen in parameter cinterval. In this example we are interested in testing all possible climate windows anywhere up to 100 days before the biological record.

`range = c(100, 0)`


**type**

There are two distinct ways to carry out a climate window analysis. In our example of chick mass we considered 'relative' climate windows, where the placement of the window will vary depending on the time of the biological response (e.g. 1-2 days *before hatching*). These relative windows assume that each individual may have the same *relative* response to climate, but the exact dates will vary between individuals.

Alternatively, there may be situations where we expect all individuals to respond to the same climatic period (e.g. average April temperature). In this case, we assume that all climate windows will have an 'absolute' date that will be taken as day 0. In this case, the additional parameter *refday* will set the day and month of day 0 for the climate window analysis.

In our example using the `Mass` dataset, we test absolute climate windows with a starting date of May 20th.

`type = "absolute"`

`refday = c(20, 5)`


**stat**

While we now have the ability to extract climate data for all possible climate windows, the aggregate statistic with which we analyse this data may also influence our results. Most commonly, we consider the mean value of climate within each window, yet it may be more appropriate to consider other possible aggregate statistics, such as maximum, minimum or slope.

The parameter stat allows users to select the aggregate statistic of their choice.

`stat = "mean"`


**func**

Although the relationship between climate and our biological response may often be linear, there is a potential for more complex relationships. The parameter func allows users to select from a range of possible relationships, including linear ("lin"), quadratic ("quad"), cubic ("cub"), logarithmic ("log") and inverse ("inv").

`func = "lin"`

### The finished product

After choosing each of our parameter values, we can finally input our choices into the `climatewin` function.

The below `climatewin` syntax will test the linear effect of mean temperature on mass across all windows between 0 and 100 days before May 20th.

*NOTE: Analysis with climwin can require large amounts of computational power. Please be patient with the results.*
```{r}
library(climwin)
```

```{r, eval = FALSE}

MassWin <- climatewin(xvar = list(Temp = MassClimate$Temp),
                      cdate = MassClimate$Date,
                      bdate = Mass$Date,
                      baseline = lm(Mass ~ 1, data = Mass),
                      cinterval = "day",
                      range = c(100, 0),
                      type = "absolute", refday = c(20, 05),
                      stat = "mean",
                      func = "lin")
                       
```

The product of our climate window analysis is a list object, here we've called it `MassWin`.

The object `MassWin` has three key components.

* Dataset is an ordered data frame summarising the results of all tested climate windows. The climate windows are ordered by delta AICc (the difference between the AICc of each model and the null model). Note, when calling the dataset we need to specify that we're interested in the dataset from the first list item (i.e. [[1]]). As we will discuss below, it is possible to test multiple combinations of parameter levels, which will output multiple list items. 

```{r, eval = FALSE}

    head(MassWin[[1]]$Dataset)

```

| deltaAICc	|WindowOpen	|WindowClose	|ModelBeta	|...     |ModWeight |...   |
|:----------|:----------|:------------|:----------|:-------|:---------|:-----|
|	-64.81496	|72	        | 15	        |-4.481257	|	...    |0.0278688	|...   |
|	-64.55352	|72	        | 14	        |-4.485161	|	...    |0.0244539	|...   |
|	-64.53157	|73	        | 15	        |-4.510254	|	...    |0.0241870	|...   | 
|	-64.40163	|73	        | 14	        |-4.517427	|	...    |0.0226655	|...   |
|	-64.30387	|72	        | 13	        |-4.500146	|	...    |0.0215842	|...   |
|	-64.20857	|73	        | 13	        |-4.533436	|	...    |0.0205799	|...   |

As we can see above, the best climate window is 72 - 15 days before May 20th, equivalent to March and April temperature.


* As the name suggests, BestModel is a model object showing the relationship between temperature and mass within the  strongest climate window.

```{r, eval = FALSE}

    MassWin[[1]]$BestModel

```

```{r, eval = FALSE}

Call:
lm(formula = Yvar ~ climate, data = modeldat)

Coefficients:
(Intercept)      climate  
    163.544       -4.481  

```


* Finally, BestModelData is a data frame containing the raw climate and biological data used to fit BestModel.

```{r, eval = FALSE}

    head(MassWin[[1]]$BestModelData)

```

|Yvar  |climate |
|:-----|:-------|
|140   |6.068966|
|138   |6.160345|
|136   |6.781034|
|135   |6.877586|
|134   |6.713793|
|134   |6.120690|

***********************

## Visualising our data

With the `climatewin` code above, we have compared all possible climate windows for our `Mass` dataset. However, one danger of the exploratory approach that we employ is that there is a risk of detecting seemingly suitable climate windows simply by statistical chance. In other words, if we fit enough climate windows one of them will eventually look good.

To overcome this concern, we have designed a number of plotting options that allow users to visualise their climate window data and determine if the 'best' window detected represents a real period of climate sensitivity or simply a statistical fluke.

We have designed 6 different plotting functions to help users visualise and interpret their `climatewin` results. Below we will discuss each plot in detail.

### Plot delta AICc values

As a first step, we can look at the distribution of delta AICc values across all tested climate windows. In the below plot, blue regions represent climate windows with limited strength (AICc values similar to the null model) while red regions show strong models.

In our `Mass` example, we can see an obvious region of strong windows around the left of the graph. This seems to suggest that there is a clear area of sensitivity.

```{r, fig.width = 4, fig.height = 4}

plotdelta(dataset = MassOutput)

```

