---
title: "New features climwin v2.0"
author: "Liam D. Bailey and Martijn van de Pol"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{climwin}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
************************************
  
## Performance

```{r}
data("bench_results")
```

```{r}
ggplot() +
  geom_line(data = bench_results,
            aes(x = window_size, y = median, colour = expression),
            linewidth = 1) +
  geom_point(data = bench_results,
            aes(x = window_size, y = median, colour = expression),
            shape = 21) +
  theme_classic()
```


************************************

## Non-linear (quadratic or cubic) climate effects

```{r}
data("Mass")
data("MassClimate")
```

**Old syntax**
```{r}
old_output <- slidingwin(xvar = list(Temp = MassClimate$Temp),
                         cdate = MassClimate$Date, 
                         bdate = Mass$Date, 
                         baseline = lm(Mass ~ 1, data = Mass),
                         range = c(150, 0), 
                         type = "absolute", refday = c(20, 5),
                         stat = "mean",
                         func = c("cub"),
                         ## Need to specify cmissing method
                         cmissing = "method1",
                         cinterval = "day")
```

**New syntax**

Specify any model structure manually.

```{r}
new_output <- run_slidingwin(range = 0:150, type = "absolute", refday = "20/05/2020",
                             climate_data = MassClimate,
                             xvar = "Temp", fn = mean,
                             cinterval = "day",
                             bio_data = Mass,
                             basemodel = lm(Mass ~ poly(climate, 3), data = bio_data))
```

```{r}
plotdelta(old_output[[1]]$Dataset)
```

```{r}
plot_delta(new_output)
```

************************************

## Monthly or weekly data

```{r}
data("Mass")
data("MassClimate")
```

**Old syntax**
```{r}
old_output <- slidingwin(xvar = list(Temp = MassClimate$Temp),
                         cdate = MassClimate$Date, 
                         bdate = Mass$Date, 
                         baseline = lm(Mass ~ 1, data = Mass),
                         range = c(70, 0), 
                         type = "absolute", refday = c(20, 5),
                         stat = "mean",
                         func = c("lin"), cmissing = FALSE,
                         cinterval = "week")
```

**New syntax**

In the new syntax we separate out data preparation and window fitting into separate functions. This has several benefits a) easier to spot mistakes & debug. You can inspect the input data before running windows b) customisable. Infinite number of ways to process data before windows, c) easier to maintain code (`run_slidingwin`) is now much leaner.

```{r}
MassClimate_week <- trans_clim_interval(MassClimate, xvar = "Temp", cdate = "Date", aggfunc = mean, cinterval = "week")
```

```{r}
new_output <- run_slidingwin(range = 0:70, type = "absolute", refday = "20/05/2020",
                             climate_data = MassClimate_week,
                             xvar = "Temp", fn = mean,
                             cinterval = "week",
                             bio_data = Mass,
                             basemodel = lm(Mass ~ climate, data = bio_data))
```

```{r}
plotdelta(old_output[[1]]$Dataset)
```

```{r}
plot_delta(new_output)
```

************************************

## Thresholds & growing degree days

```{r}
data("Mass")
data("MassClimate")
```

**Old syntax**
```{r}
old_output <- slidingwin(xvar = list(Temp = MassClimate$Temp),
                         cdate = MassClimate$Date, 
                         bdate = Mass$Date, 
                         baseline = lm(Mass ~ 1, data = Mass),
                         range = c(12, 0), 
                         type = "absolute", refday = c(20, 5),
                         stat = "sum",
                         func = c("lin"), cmissing = FALSE,
                         cinterval = "month", upper = 10, binary = TRUE)
```

**New syntax**

In the new syntax we separate out data preparation and window fitting into separate functions. This has several benefits a) easier to spot mistakes & debug. You can inspect the input data before running windows b) customisable. Infinite number of ways to process data before windows, c) easier to maintain code (`run_slidingwin`) is now much leaner.

```{r}
## Convert to months (NOTE: If you want to apply threshold to data BEFORE taking monthly summary, just reverse the order!!)
MassClimate_mo <- trans_clim_interval(MassClimate, xvar = "Temp", cdate = "Date", aggfunc = mean, cinterval = "month")
MassClimate_mo
```

```{r}
## Create a new binary column
MassClimate_mo_binary <- append_clim_threshold(MassClimate_mo, xvar = "Temp", upper = 10, binary = TRUE)
MassClimate_mo_binary
```

```{r}
new_output <- run_slidingwin(range = 0:12, type = "absolute", refday = "20/05/2020",
                             climate_data = MassClimate_mo_binary,
                             xvar = "threshold", fn = sum,
                             cinterval = "month",
                             bio_data = Mass,
                             basemodel = lm(Mass ~ climate, data = bio_data))
```

```{r}
plotdelta(old_output[[1]]$Dataset)
```

```{r}
plot_delta(new_output)
```

************************************

## Incomplete climate data

```{r}
data("Mass")
data("MassClimate")
MassClimate_gaps <- MassClimate |> 
  mutate(Date = lubridate::dmy(Date)) |> 
  ## Remove every record in 1st of the month!
  dplyr::filter(lubridate::day(Date) != 1)
```

**Old syntax**
```{r}
## This fails
# old_output <- slidingwin(xvar = list(Temp = MassClimate_gaps$Temp),
#                          cdate = MassClimate_gaps$Date, 
#                          bdate = Mass$Date, 
#                          baseline = lm(Mass ~ 1, data = Mass),
#                          range = c(150, 0), 
#                          type = "absolute", refday = c(20, 5),
#                          stat = "mean",
#                          func = c("lin"), cmissing = FALSE,
#                          cinterval = "day")

old_output <- slidingwin(xvar = list(Temp = MassClimate_gaps$Temp),
                         cdate = MassClimate_gaps$Date, 
                         bdate = Mass$Date, 
                         baseline = lm(Mass ~ 1, data = Mass),
                         range = c(150, 0), 
                         type = "absolute", refday = c(20, 5),
                         stat = "mean",
                         func = c("lin"),
                         ## Need to specify cmissing method
                         cmissing = "method1",
                         cinterval = "day")
```

**New syntax**

Fix gaps *BEFORE*

```{r}
MassClimate_fixed <- repair_climate(MassClimate_gaps, cdate = "Date", xvar = "Temp")
```

```{r}
new_output <- run_slidingwin(range = 0:150, type = "absolute", refday = "20/05/2020",
                             climate_data = MassClimate_fixed,
                             xvar = "Temp", fn = mean,
                             cinterval = "day",
                             bio_data = Mass,
                             basemodel = lm(Mass ~ climate, data = bio_data))
```

```{r}
plotdelta(old_output[[1]]$Dataset)
```

```{r}
plot_delta(new_output)
```

